\LARGE \bf \pyRobots{}, a toolset for robot executive control

### 2014-09-18 paper.tex
Completed studies section

Actions can also call each other, allowing natural decomposition of complex
tasks into sub-actions and atomic actions.

@@

The development of \pyRobots{} started in 2011, driven by the requirements of a
18-min long theatre performance between a human actor and a robot. Since that time, the toolkit has
iteratively matured, and has been deployed and used on several projects and
platforms. We first briefly recap these deployments with references to the
corresponding publications, and then present our latest study, in a nursery,
that recently acted as a stress-test for the library.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
\begin{center}
\begin{tabular}{p{0.9\columnwidth}}
    \toprule
    {\bf Manipulation} \\
     {\tt attachobject}, {\tt basicgive}, {\tt basictake}, {\tt close\_gripper}, {\tt configure\_grippers}, {\tt grab\_gripper}, {\tt handover}, {\tt hide}, {\tt open\_gripper}, {\tt pick}, {\tt put}, {\tt put\_accessible}, {\tt release\_gripper}, {\tt show}, {\tt take} \\
     \midrule
    {\bf Gaze control} \\
     {\tt glance\_to}, {\tt look\_at}, {\tt sweep}, {\tt switch\_active\_stereo\_pair}, {\tt track}, {\tt cancel\_track} \\
     \midrule
    {\bf Navigation} \\
     {\tt carry}, {\tt follow}, {\tt cancel\_follow}, {\tt goto}, {\tt moveclose}, {\tt waypoints} \\
     \midrule
    {\bf Local navigation} \\
     {\tt dock}, {\tt rotate}, {\tt translate} \\
     \midrule
    {\bf Posture configuration} \\
     {\tt extractpose}, {\tt idle}, {\tt manipose}, {\tt movearm}, {\tt rest}, {\tt setpose}, {\tt settorso}, {\tt tuckedpose} \\
     \bottomrule
\end{tabular}
\end{center}
\caption{Examples of \pyRobots{} actions, mostly extracted from experiments involving
    the PR2 robot.}

@@

\subsubsection{Roboscopie} The first complex deployment scenario for \pyRobots{} was
indeed a theatre performance entitled
\emph{Roboscopie}~\cite{lemaignan2012roboscopie}. The toolkit was created out of
the need to control a complex robot (a Willow Garage PR2) in a structured way,
as required for a live performance with a human\footnote{The original \pyRobots{}
script used for the performance can be accessed
from~\url{http://www.laas.fr/roboscopie}. Note that this script does not
reflect anymore the toolkit usage.}

A total of 47 \pyRobots{} actions were written for this event
(Table~\ref{pyrobots_actions} gives a few examples), and the toolkit proved to
be especially convenient to interleave calls to both ROS-based and {\sc
GenoM}-based\footnote{{\sc GenoM} is software component as well as a middleware
developed at LAAS-CNRS} components through their respective Python bindings.

\begin{itemize}
    \item PR2, with ROS and pocolibs
    \item Nao, with ROS and NAOqi
    \item Ranger with Aseba and ROS
\end{itemize}

\begin{itemize}
    \item roboscopie
    \item LAAS architecture
    \item CoWriter
    \item Ranger
\end{itemize}

@@

\subsubsection{Stress-test in a nursery}
\label{croq}

We organized in June 2014 a pilot study in a kindergarten (on campus) to
stress-test the hardware and software of the Ranger
robot~\cite{mondada2014ranger}. The Ranger robot is an autonomous robot designed
for interaction with children. As it was primarily designed to support children
with tidying up, it is shaped as a wooden ``box on wheels''
(Figure~\ref{expe-nursery}). Hardware-wise, it features an RGB-D camera, IR
sensors, a range and bearing sensor, a physical front bumper, a scale, a
removable pacifier (used as a soft on/off switch by the children) and it is
covered with capacitive tactile panels. The wooden panels cover 186 RGB LEDs,
enabling various light patterns, it can play sounds, the two eyes can be fully
controlled (2 DoFs pupils and actuated eye-lids). Wheel controllers are designed
such as they switch to freewheel as soon as the robot is pushed by the children.
The robot is powered by an embedded Gumstix Overo (ARM7l) board running a custom
Linux image, assisted by three custom microcontroller boards (for low-level
hardware interfacing).

@@

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{ranger-side}
        \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
        The behaviour of the robot is controlled with \pyRobots{}.}
        \label{expe-nursery}
\end{figure}

We brought two Ranger robots to a kindergarten (17 children, age: M=3.88,
SD=0.65), and after a brief introduction (showing them that removing the
pacifier would ``wake up'' the robot, and putting it back would make it ``fall
asleep''), we invited the children to freely play with the robots and integrate
them into their usual games. The experiment lasted 68 minutes.

We purposefully designed a simple behavioural scheme
(Listing~\ref{croquignole_no_move}): two background actions
(\python|background_blink| and \python|look_at_touches|) run continuously, and 4
events are monitored (pacifier added/removed, toy added and bumper hit).

@@

\begin{pythoncode}
    @action
    def on_pacifier(robot):
        robot.look_at_lolette()
        robot.blink()
        sleep = robot.fall_asleep()
        robot.lightbar(ramp=colors.RAINBOW).wait()
        sleep.wait()

@@

        while True:
            time.sleep(0.1)
\end{pythoncode}
\caption{Source of the high-level behaviours running on the robots during the
nursery pilot (some behaviours like battery management have been omitted for
clarity).}
\label{croquignole_no_move}
\end{listing}

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{croquignole-expe}
    \caption{Events fired and actions started per minute, over the 68 min-long
    experiment. The robot sustains an average of 70 actions triggered by minute over
    more than an hour, with peaks at 200 actions/minute.}

\end{figure}

>>>>

Actions can also call each other (nested actions), allowing natural
decomposition of complex tasks into sub-actions and atomic actions (we note
hereafter $depth$ the nesting level of an action, \ie the maximum number of
calls to sub-actions until an atomic action is reached).

@@

The development of \pyRobots{} started in 2011, driven by the requirements of a
18min long theatre performance between a human actor and a robot. Since that
time, the toolkit has iteratively matured, and has been deployed and used on
several projects and platforms. We first briefly recap these deployments with
references to the corresponding publications, and then present our latest study,
in a kindergarten, that recently acted as a stress-test for the library.

\begin{table}[h]
    \centering
    \begin{tabular}{lll}
        \toprule
        Study                 & Robot & Middleware \\ \midrule
        Roboscopie            & PR2       & ROS, {\sc GenoM} \\
        Interactive Grounding & PR2, LAAS-CNRS's Jido & ROS, {\sc GenoM} \\
        CoWriter              & Nao       & ROS, NAOqi  \\
        Ranger at kindergarten & EPFL's Ranger & ROS, aseba  \\
        \bottomrule
    \end{tabular}
    \caption{Overview of significant experiments and studies where \pyRobots{}
    has been used.}
\end{table}

\subsubsection{Roboscopie} The first complex deployment scenario for \pyRobots{} was
a theatre performance entitled \emph{Roboscopie}~\cite{lemaignan2012roboscopie}.
The toolkit was created out of the need to control a complex robot (a Willow
Garage PR2) in a structured way, as required for a live performance with a
human\footnote{The original \pyRobots{} script used for the performance can be
accessed from~\url{http://www.laas.fr/roboscopie}. Note that this script does
not reflect anymore the toolkit usage.}

A total of 47 \pyRobots{} actions were written for this event
(Table~\ref{pyrobots_actions} gives a few significant examples), and the toolkit
proved to be especially convenient to interleave calls to both ROS-based and
{\sc GenoM}-based~\cite{mallet2010genom3} components through their respective
Python bindings.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[ht!]
\begin{center}
\begin{tabular}{p{0.9\columnwidth}}
    \toprule
    {\bf Manipulation} \\
     {\tt attachobject}, {\tt basicgive}, {\tt basictake}, {\tt close\_gripper}, {\tt configure\_grippers}, {\tt grab\_gripper}, {\tt handover}, {\tt hide}, {\tt open\_gripper}, {\tt pick}, {\tt put}, {\tt put\_accessible}, {\tt release\_gripper}, {\tt show}, {\tt take} \\
     \midrule
    {\bf Gaze control} \\
     {\tt glance\_to}, {\tt look\_at}, {\tt sweep}, {\tt switch\_active\_stereo\_pair}, {\tt track}, {\tt cancel\_track} \\
     \midrule
    {\bf Navigation} \\
     {\tt carry}, {\tt follow}, {\tt cancel\_follow}, {\tt goto}, {\tt moveclose}, {\tt waypoints} \\
     \midrule
    {\bf Local navigation} \\
     {\tt dock}, {\tt rotate}, {\tt translate} \\
     \midrule
    {\bf Posture configuration} \\
     {\tt extractpose}, {\tt idle}, {\tt manipose}, {\tt movearm}, {\tt rest}, {\tt setpose}, {\tt settorso}, {\tt tuckedpose} \\
     \bottomrule
\end{tabular}
\end{center}
\caption{Examples of \pyRobots{} actions (extracted from experiments involving
the PR2 robot).}

@@

\subsubsection{Interactive Grounding} This study on perspective-taking and grounding
in human-robot interaction~\cite{lemaignan2013talking} is one of the several
experiments that was built on top of the LAAS-CNRS cognitive
architecture~\cite{lemaignan2014human}. In this large architecture, \pyRobots{}
is one of the two available executive controllers, and integrates with a task
planner, a symbolic knowledge base, several motion planner and a natural
language processor. \pyRobots{} was favoured over the second execution
controller (based on PRS) for smaller studies, as a more powerful alternative to
bare \textit{ad-hoc} scripts.

\subsubsection{CoWriter} The {\sc CoWriter} project~\cite{hood2015those}
(Figure~\ref{expe-cowriter}) involves an Aldebaran Nao robot and young students
in a \emph{learning by teaching} interaction: the children have to teach the
robot how to handwrite. This experiment involved a dual ROS/NAOqi environment,
bridged within \pyRobots{}. The high-level behaviour was implemented as a finite
state machine ({\sc SMACH}) whose state corresponded to \pyRobots{} actions.
While the execution controller was subsequently simplified (towards a tighter
interaction scenario), this illustrates how \pyRobots{} -- relying on the
ubiquitous Python language -- can possibly blend into and interact with existing
tools.

@@

\subsubsection{Stress-test in a kindergarten}
\label{croq}

We organized in June 2014 a pilot study in a kindergarten (on campus) to
stress-test the hardware and software of the Ranger
robot~\cite{mondada2014ranger}. The Ranger robot is an autonomous robot designed
for interaction with children. As it was primarily designed to support children
with tidying up, it is shaped as a wooden ``box on wheels''
(Figure~\ref{expe-nursery}). Hardware-wise, it features an RGB-D camera, IR
sensors, a range and bearing sensor, a physical front bumper, a scale, a
removable pacifier (used as a soft on/off switch by the children) and it is
covered with capacitive tactile panels. The wooden panels hide 186 RGB LEDs,
enabling various light patterns, it can play sounds, the two eyes can be fully
controlled (2 DoFs pupils and actuated eye-lids). Wheel controllers are designed
such as they switch to freewheel as soon as the robot is pushed by the children.
The robot is powered by an embedded Gumstix Overo board (ARM7l) running a custom
Linux image, assisted by three custom microcontroller boards (for low-level
hardware interfacing).

@@

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{ranger-side}
        \caption{Infants playing freely (!) with an EPFL's \emph{Ranger} robot.
        The behaviours of the robot are implemented via \pyRobots{}.}
        \label{expe-nursery}
\end{figure}

We brought two such Ranger robots to a kindergarten (17 children, age: M=3.88,
SD=0.65), and after a brief introduction (showing them that removing the
pacifier would ``wake up'' the robot, and putting it back would make it ``fall
asleep''), we invited the children to freely play with the robots and integrate
them into their usual games. The experiment lasted 68 minutes.

We purposefully designed a simple behavioural scheme
(Listing~\ref{croquignole_no_move}): two background actions
(\python|background_blink| and \python|look_at_touches|) run continuously, and 5
events are monitored (pacifier added/removed, toy added/removed and bumper hit).

@@

\begin{pythoncode}
    @action
    def on_pacifier(robot):
        robot.look_at_pacifier()
        robot.blink()
        sleep = robot.fall_asleep()
        robot.lightbar(ramp=colors.RAINBOW).wait()
        sleep.wait()

@@

        while True:
            time.sleep(0.1)
\end{pythoncode}
\caption{Simplified source of the high-level behaviours running on the robots during the
kindergarten study (some behaviours like battery management have been omitted for
clarity).}
\label{croquignole_no_move}
\end{listing}

Figure~\ref{croq_actions} shows the average number of events fired and action
(including sub-actions) \emph{started} per minute, for the whole experiment. The
diagram reveals several peaks above 180 actions/minutes with a global average of
70 actions/minutes, which was surprisingly high considering the simple
behavioural scheme we proposed and the limited actions' depth
($depth_{max}=depth_{\tt look\_at\_touches}=3$) . As a safety measure, we also
set for this study a hard limit to the maximum amount of actions allowed to run
in parallel (20 actions): again, to our surprise, this limit was hit 31 times
during the experiment.

\begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{croquignole-expe}
    \caption{Events fired and actions started per minute, over the 68min long
    experiment. The robot sustains an average of 70 actions triggered by minute over
    more than an hour, with peaks at 200 actions/minute.}
    \label{croq_actions}
\end{figure}

The results of this stress-test provide a reasonable baseline for the
performances in terms of parallelism and debug/introspection capabilities that
can be expected from an executive controller.
